Trends of SDR in Thessaloniki

### 2023-07-09 Article.Rmd
Commit 2023-07-08 21:10
Commit 2023-07-09 20:14

## create local citations file
source("~/CODE/FUNCTIONS/R/gather_citations.R")

@@

We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table\ \ref{tab:trendtable}).
In this time period, we could not find any significant break or change in the variability of the trends.
Other studies for the European region are reporting a change of the SDR slope, around the '80, few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].

@@

Cumulative sum [@Regier2019]

```{r cumsum, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics(c(
    "./images/CumulativeDailyCumSum-1.png",    # ALL sky
    "./images/CumulativeDailyCumSum-5.png",    # CLEAR sky
    "./images/CumulativeDailyCumSum-9.png"))   # Cloud sky
```

more about trend in relation to other trends

@@

\footnotesize
```{r trendtable, echo=F}
LT <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE))
LT <- LT[grep("SDR", LT$var), ]
LT$slope.sd           <- NULL
LT$slope.ConfInt_0.95 <- NULL
LT$slope.ConfInt_0.99 <- NULL
LT$var                <- NULL
setorder(LT, -slope.stat_sig)
rownames(LT) <- NULL
names(LT)    <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
## no need all are 100%
LT$`Trend statistical signif. [%]` <- NULL

panderOptions("table.alignment.default", "center")
pander(LT,
       caption = "\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions. All cases have a statistical significance above $99.9\\%$.")
# knitr::kable(LT, caption = "Trends of daily means by sky conditions.")

## some vars to use later
alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend [%/year]`], 2)
cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.", `Trend [%/year]`], 2)
cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend [%/year]`], 2)
```
\normalsize

>>>>

## create local citations file and links here
source("~/CODE/FUNCTIONS/R/gather_citations.R")

@@

We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table\ \ref{tab:trendtable}).
We present only the plot for the trend under all-sky conditions (Figure\ \ref{fig:trendALL}), as the plots for clear-sky condition and cloud-sky condition, are very similar.
In the studied time period, there is no significant break or change in the variability of the overall trends.
Other studies for the European region are reporting a change of the SDR slope, around the 1980, few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].

<!-- \footnotesize -->
```{r trendtable, echo=F}
LT <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE))
LT <- LT[grep("SDR", LT$var), ]
LT$slope.sd           <- NULL
LT$slope.ConfInt_0.95 <- NULL
LT$slope.ConfInt_0.99 <- NULL
LT$var                <- NULL
setorder(LT, -slope.stat_sig)
rownames(LT) <- NULL
names(LT)    <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
## no need all are 100%
LT$`Trend statistical signif. [%]` <- NULL

panderOptions("table.alignment.default", "center")
pander(LT,
       caption = paste0("\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ". For all cases the statistical significance is above $99.9\\%$."))
# knitr::kable(LT, caption = "Trends of daily means by sky conditions.")
## some vars to use later
alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend [%/year]`], 2)
cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.", `Trend [%/year]`], 2)
cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend [%/year]`], 2)
```
<!-- \normalsize -->

Although, using the cumulative sums plots we can reveal some structure within the trends and some difference between the sky conditions [@Regier2019].

Comparining the conditions

Variability issues

```{r cumsum, echo = FALSE, fig.cap='Running plots of the cumulative sum of the daily SDR for three different sky conditions', fig.subcap=c('All-sky conditions', 'Clear-sky conditions', 'Cloud-sky conditions'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
knitr::include_graphics(c(
    "./images/CumulativeDailyCumSum-1.png",    # ALL sky
    "./images/CumulativeDailyCumSum-5.png",    # CLEAR sky
    "./images/CumulativeDailyCumSum-9.png"))   # Cloud sky
```

more about trend in relation to other trends
Aerosols and clouds

@@

@@
